Capstone Movielens Report

### 2025-05-07 reports/capstone-movielens-report.draft4.tex

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-06}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  mu.RMSE }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{mean}\NormalTok{(MSEs))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu.RMSE}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.060346
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{deviation }\OtherTok{\textless{}{-}} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{6}\NormalTok{, }\FloatTok{0.1}\NormalTok{) }\SpecialCharTok{{-}} \DecValTok{3}
\FunctionTok{print}\NormalTok{(deviation)}

\NormalTok{rmse\_value }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(deviation, }\ControlFlowTok{function}\NormalTok{(diff)\{}
  \FunctionTok{rmse}\NormalTok{(test\_set}\SpecialCharTok{$}\NormalTok{rating }\SpecialCharTok{{-}}\NormalTok{ mu }\SpecialCharTok{+}\NormalTok{ diff)}
\NormalTok{\})}

\FunctionTok{plot}\NormalTok{(deviation, rmse\_value, }\AttributeTok{type =} \StringTok{"l"}\NormalTok{)}

\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Minimum RMSE is achieved when the deviation from the mean is: \%s"}\NormalTok{, }
\NormalTok{        deviation[}\FunctionTok{which.min}\NormalTok{(rmse\_value)])}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-42-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-54-1.pdf}

>>>>

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-07}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu.MSEs }\OtherTok{\textless{}{-}} \FunctionTok{naive\_model\_MSEs}\NormalTok{(mu)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
  \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{fold\_No =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{MSE =}\NormalTok{ mu.MSEs) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{data.plot}\NormalTok{(}\AttributeTok{title =} \StringTok{"MSE resuls of the 5{-}fold CV method applied to the Overall Mean Rating Model"}\NormalTok{,}
                \AttributeTok{xname =} \StringTok{"fold\_No"}\NormalTok{, }
                \AttributeTok{yname =} \StringTok{"MSE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-40-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu.RMSE }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{mean}\NormalTok{(mu.MSEs))}
\NormalTok{mu.RMSE}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.061191
\end{verbatim}

\begin{noteblock}
For the \emph{Mean Squared Error} data visualization we used
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R\#L630}{\texttt{data.plot}
function} defined in
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R\#L484}{\emph{Model
Data Visualization} section} of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R}{common-helper.functions.R}
script.

Here we also used
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R\#L54}{\texttt{naive\_model\_MSEs}
function} defined in the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R}{same
script} to compute \emph{Mean Squared Errors} using \emph{5-Fold Cross
Validation} method:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{naive\_model\_MSEs }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(val) \{}
  \FunctionTok{sapply}\NormalTok{(edx\_CV, }\ControlFlowTok{function}\NormalTok{(cv\_item)\{}
    \FunctionTok{mse}\NormalTok{(cv\_item}\SpecialCharTok{$}\NormalTok{validation\_set}\SpecialCharTok{$}\NormalTok{rating }\SpecialCharTok{{-}}\NormalTok{ val)}
\NormalTok{  \})}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

One more function we will need for further analysis of the current model
is the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R\#L59}{\texttt{naive\_model\_RMSE}
one} defined in the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R}{common-helper.functions.R}
script already mentioned above:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{naive\_model\_RMSE }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(val)\{}
  \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{mean}\NormalTok{(}\FunctionTok{naive\_model\_MSEs}\NormalTok{(val)))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\paragraph{\texorpdfstring{Ensure that \texttt{mu.RMSE} value is the
best for the current
model}{Ensure that mu.RMSE value is the best for the current model}}\label{ensure-that-mu.rmse-value-is-the-best-for-the-current-model}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  deviation }\OtherTok{\textless{}{-}} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{6}\NormalTok{, }\FloatTok{0.1}\NormalTok{) }\SpecialCharTok{{-}} \DecValTok{3}

\NormalTok{  deviation.RMSE }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(deviation, }\ControlFlowTok{function}\NormalTok{(delta)\{}
    \FunctionTok{naive\_model\_RMSE}\NormalTok{(mu }\SpecialCharTok{+}\NormalTok{ delta)}
\NormalTok{  \})}
  \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"RMSE values have been computed for \%1 deviations from the Overall Mean Rating."}\NormalTok{,}
           \FunctionTok{length}\NormalTok{(deviation))}
\end{Highlighting}
\end{Shaded}

Let's make a quick investigation of the \texttt{deviation.RMSE} result
we have just got:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data.frame}\NormalTok{(}\AttributeTok{delta =}\NormalTok{ deviation, }
           \AttributeTok{delta.RMSE =}\NormalTok{ deviation.RMSE) }\SpecialCharTok{|\textgreater{}} 
\FunctionTok{data.plot}\NormalTok{(}\AttributeTok{title =} \FunctionTok{TeX}\NormalTok{(r}\StringTok{\textquotesingle{}[RMSE as a function of deviation ($\textbackslash{}delta$) from the Overall Mean Rating ($\textbackslash{}hat\{mu\}$)]\textquotesingle{}}\NormalTok{),}
              \AttributeTok{xname =} \StringTok{"delta"}\NormalTok{, }
              \AttributeTok{yname =} \StringTok{"delta.RMSE"}\NormalTok{, }
              \AttributeTok{xlabel =} \FunctionTok{TeX}\NormalTok{(r}\StringTok{\textquotesingle{}[$\textbackslash{}delta$]\textquotesingle{}}\NormalTok{), }
              \AttributeTok{ylabel =} \StringTok{"RMSE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in is.na(xlabel): is.na() applied to non-(list or vector) of type 'expression'
\end{verbatim}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-44-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{which\_min\_deviation }\OtherTok{\textless{}{-}}\NormalTok{ deviation[}\FunctionTok{which.min}\NormalTok{(deviation.RMSE)]}
\NormalTok{min\_rmse }\OtherTok{=} \FunctionTok{min}\NormalTok{(deviation.RMSE)}

\FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Minimum RMSE is achieved when the deviation from the mean is: \%1"}\NormalTok{,}
\NormalTok{         which\_min\_deviation)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Minimum RMSE is achieved when the deviation from the mean is: 0
\end{verbatim}

\begin{verbatim}
## Log is not open.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Is the previously computed RMSE the best for the current model: \%1"}\NormalTok{,}
\NormalTok{         mu.RMSE }\SpecialCharTok{==}\NormalTok{ min\_rmse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Is the previously computed RMSE the best for the current model: FALSE
\end{verbatim}

\begin{verbatim}
## Log is not open.
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-57-1.pdf}
